E4-analysis-replication-in-lab

### 2023-07-06 group-d/E4-analysis-replication-in-lab.Rmd
added more merced data and did first pass analysis

```{r , include=FALSE}
library(papaja)
library(jsonlite)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
```

@@

```{r E4-in-lab-load-data}
# data.files <- list.files('data/run-2', full.names = TRUE)
# data.tables <- lapply(data.files, function(file){
#   data.table <- fromJSON(file)
#   return(data.table)
# })
data.tables = readRDS("data/in-lab-merced/verb_bias_ET_data_10_13_2022.rds")
all.data <- bind_rows(data.tables)
```

@@

```{r E4-filter-data-for-replication-analysis}
task.data <- all.data %>%
  dplyr::filter(compatibility != 'NA', compatibility != 'filler') %>%
  dplyr::select(user_id, trial_index, rt, images, webgazer_data, mouse_events, compatibility, audio, target_instrument, target_animal, webgazer_targets)

@@

```{r E4-categorizing-first-mouse-moves}
#First, create a data set with each trial, adding column for which object the mouse moves over first.
# WTF IS HAPPENING with TRIAL ID?? 
first.move <- mousetracking.data %>%
  select(user_id, trialID, object, type, compatibility, target_instrument, target_animal, images) %>%
  group_by(user_id, trialID) %>%
  filter(type=="enter") %>%
  slice(1) %>%
  rowwise() %>%
  mutate(which_loc = which(images == object)-1) %>%
  ungroup() %>%
  mutate(instrument_loc = str_sub(target_instrument, start=-1, end=-1)) %>%
  mutate(animal_loc = str_sub(target_animal, start=-1, end=-1)) %>%
  mutate(first.move.type = case_when(
    which_loc == instrument_loc ~ 'instrument',
    which_loc == animal_loc ~ 'animal',
    TRUE ~ 'other')) %>%
  mutate(is.mouse.instrument = if_else(first.move.type == 'instrument', 1, 0)) %>%
  mutate(compatibility = factor(compatibility))
```

```{r E4-summarizing-first-mouse-moves-for-plotting}
# Summarize the data by subject, calculating proportion of trials the the first move was to the animal, instrument, or other.
first.move.subject.summary <- first.move %>%
  group_by(user_id, compatibility) %>%
  summarize(prop.animal = mean(first.move.type == 'animal'),
            prop.instrument = mean(first.move.type == 'instrument'),
            prop.other = mean(first.move.type == 'other')) %>%
  pivot_longer(c('prop.animal', 'prop.instrument', 'prop.other'), names_to="target_type", values_to="proportion")

@@

```{r E4-mouse-moves-analysis}
contrasts(first.move$compatibility) <- cbind(c(1/3, -2/3, 1/3), c(-1/2, 0, 1/2))

E4_mouse_moves_model <- glmer(is.mouse.instrument ~ compatibility + (1 + compatibility | subject) + (1 | trialID), data=first.move, family="binomial",
                              glmerControl(optimizer = "bobyqa"))

@@

```{r E4-summarizing-gaze-data-for-plotting}
#Summarize data for plotting
eyetracking.figure.2.data <- eyetracking.data.with.time.windows %>%
  filter(between(t.window , -200, 4000)) %>%
  group_by(user_id, compatibility, t.window) %>%
  summarize(p.animal = mean(in.roi.animal), p.instrument = mean(in.roi.instrument)) %>%
  pivot_longer(c('p.animal', 'p.instrument'), names_to="object_type", values_to="prop_fixations") %>%
  #mutate(prop_fixations = if_else(is.na(prop_fixations), 0, prop_fixations)) %>%
  group_by(compatibility, t.window, object_type) %>%
  summarize(M=mean(prop_fixations), SE=sd(prop_fixations)/sqrt(n()))
```

@@

fig2<-ggplot(eyetracking.figure.2.data %>% 
  mutate(compatibility = factor(compatibility, 
                                levels = c("modifier", "equibiased", "instrument" ), 
                                labels = c("Modifier", "Equi-biased", "Instrument" ))), 
  aes(x=t.window, y=M, ymin=M-SE, ymax=M+SE, color=compatibility, fill=compatibility, linetype=object_type))+
  geom_ribbon(color=NA, alpha=0.3)+
  geom_line(size=1)+
  scale_color_manual(values = c( "#377eb8", "#e41a1c","#4daf4a"))+
  scale_fill_manual(values = c("#377eb8", "#e41a1c", "#4daf4a"))+
  scale_linetype(labels = c("Animal", "Instrument") )+
  theme_classic() +
  geom_vline(xintercept = animal.onset + 200) + 
  geom_vline(xintercept = instrument.onset + 200)+
  labs(y = "Proportion of looks", x = "Time relative to verb onset (ms)")+
  guides(color = guide_legend("Verb bias"),fill = guide_legend("Verb bias"), linetype = guide_legend("Gaze location"))
fig2
```
```{r, eval=FALSE}
saveRDS(fig2, "output/ETfig.rds")
ggsave("output/ETfig.png", plot = fig2, height = 4, width = 6)
```

```{r E4-compute-proportion-gaze-data-by-trial}
#Summarize fixations on target and instrument
eyetracking.window.summary.by.trial <- eyetracking.data.with.time.windows %>%
  group_by(subject, trialID, sound, compatibility, time.window) %>%
  summarize(prop.fixations.animal = sum(in.roi.animal) / n(),
            prop.fixations.instrument = sum(in.roi.instrument) / n()) %>%
  mutate(compatibility = factor(compatibility))
```

@@

model.time.window.1 <- lmer(prop.fixations.animal ~ compatibility + (1  | subject) + (1 | trialID), data=data.time.window.1,
                            control = lmerControl(optimizer = "bobyqa",
                                                  optCtrl = list(maxfun = 2e6)))

@@

model.time.window.2 <- lmer(prop.fixations.animal ~ compatibility + (1  | subject) + (1 | trialID), data=data.time.window.2)
```

@@

model.time.window.3 <- lmer(prop.fixations.animal ~ compatibility + (1 | subject) + (1 | trialID), data=data.time.window.3)
```
```{r E4-ET-analysis-tables}
E4_ET_model1_tab = broom.mixed::tidy(model.time.window.1) 
E4_ET_model2_tab = broom.mixed::tidy(model.time.window.2) 
E4_ET_model3_tab = broom.mixed::tidy(model.time.window.3)

>>>>

```{r , include=FALSE}
library(papaja)
library(jsonlite)
library(purrr)
#library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
```

@@

```{r E4-in-lab-load-data}
user.data.files <- list.files('data/in-lab-merced', pattern = "pushkin_users_*",full.names = TRUE)

all.user.data <- map_dfr(user.data.files, readRDS)

data.files <- list.files('data/in-lab-merced', pattern = "lex_*",full.names = TRUE)
# data.tables <- lapply(data.files, function(file){
#   data.table <- fromJSON(file)
#   return(data.table)
# })

#data.tables = readRDS("data/in-lab-merced/lex_stimulusResponses_2023-05-01.rds")
all.data <- map_dfr(data.files, readRDS) %>% 
  bind_rows(readRDS("data/in-lab-merced/verb_bias_ET_data_10_13_2022.rds")) %>% 
  arrange(user_id, trial_index, time_elapsed) %>% 
  group_by(user_id, trial_index) %>% 
  mutate(subjnum = 1:n(),
         unique_id = str_c(subjnum,"__", user_id)) %>% 
  ungroup() %>% 
  arrange(unique_id, trial_index)

*TO DO:* cross ref with angela's list of participant times/dates

@@

```{r E4-filter-data-for-replication-analysis}
task.data <- all.data %>%
  dplyr::filter(compatibility != 'NA', compatibility != 'filler') %>%
  dplyr::select(unique_id, trial_index, rt, images, webgazer_data, mouse_events, compatibility, audio, target_instrument, target_animal, webgazer_targets)

@@

```{r E4-categorizing-first-mouse-moves}
#First, create a data set with each trial, adding column for which object the mouse moves over first.
# WTF IS HAPPENING with TRIAL ID?? 
first.move <- mousetracking.data %>%
  select(unique_id, trialID, object, type, compatibility, target_instrument, target_animal, images, audio) %>%
  group_by(unique_id, trialID) %>%
  filter(type=="enter") %>%
  slice(1) %>%
  rowwise() %>%
  mutate(which_loc = which(images == object)-1) %>%
  ungroup() %>%
  mutate(instrument_loc = str_sub(target_instrument, start=-1, end=-1)) %>%
  mutate(animal_loc = str_sub(target_animal, start=-1, end=-1)) %>%
  mutate(first.move.type = case_when(
    which_loc == instrument_loc ~ 'instrument',
    which_loc == animal_loc ~ 'animal',
    TRUE ~ 'other')) %>%
  mutate(is.mouse.instrument = if_else(first.move.type == 'instrument', 1, 0)) %>%
  mutate(compatibility = factor(compatibility))
```

```{r E4-summarizing-first-mouse-moves-for-plotting}
# Summarize the data by subject, calculating proportion of trials the the first move was to the animal, instrument, or other.
first.move.subject.summary <- first.move %>%
  group_by(unique_id, compatibility) %>%
  summarize(prop.animal = mean(first.move.type == 'animal'),
            prop.instrument = mean(first.move.type == 'instrument'),
            prop.other = mean(first.move.type == 'other'),
            n_trials= n()) %>%
  filter(n_trials > 9) %>% ## CAHNGE WHEN THERE'S A BETTER WAY TO CROSSREF
  pivot_longer(c('prop.animal', 'prop.instrument', 'prop.other'), names_to="target_type", values_to="proportion")

@@

```{r E4-mouse-moves-analysis-lab}
contrasts(first.move$compatibility) <- cbind(c(1/3, -2/3, 1/3), c(-1/2, 0, 1/2))

E4_mouse_moves_model <- glmer(is.mouse.instrument ~ compatibility + (1 + compatibility | unique_id) + (1 | trialID), 
                              data=first.move, family="binomial",
                              glmerControl(optimizer = "bobyqa"))

@@

```{r E4-summarizing-gaze-data-for-plotting, eval = FALSE, warning = FALSE}
#Summarize data for plotting
eyetracking.data.with.time.windows %>%
  filter(between(t.window , -200, 4000)) %>%
  filter(unique_id %in% unique(first.move.subject.summary$unique_id)) %>% #CHANGE WHEN CROSSREF
  group_by(unique_id, compatibility, t.window) %>%
  summarize(p.animal = mean(in.roi.animal), p.instrument = mean(in.roi.instrument)) %>%
  pivot_longer(c('p.animal', 'p.instrument'), names_to="object_type", values_to="prop_fixations") %>%
  ggplot()+
  stat_summary(aes(x = t.window, y=prop_fixations,  color = object_type), fun="mean")+
  facet_wrap(~unique_id)
  #mutate(prop_fixations = if_else(is.na(prop_fixations), 0, prop_fixations)) %>%
 # group_by(compatibility, t.window, object_type) %>%
 # summarize(M=mean(prop_fixations), SE=sd(prop_fixations)/sqrt(n()))
```

```{r E4-summarizing-gaze-data-for-plotting}
#Summarize data for plotting
eyetracking.figure.2.data <- eyetracking.data.with.time.windows %>%
  filter(between(t.window , -200, 4000)) %>%
  filter(unique_id %in% unique(first.move.subject.summary$unique_id)) %>% #CHANGE WHEN CROSSREF
  group_by(unique_id, compatibility, t.window) %>%
  summarize(p.animal = mean(in.roi.animal), p.instrument = mean(in.roi.instrument)) %>%
  pivot_longer(c('p.animal', 'p.instrument'), names_to="object_type", values_to="prop_fixations") %>%
  #mutate(prop_fixations = if_else(is.na(prop_fixations), 0, prop_fixations)) %>%
  group_by(compatibility, t.window, object_type) %>%
  summarize(M=mean(prop_fixations), SE=sd(prop_fixations)/sqrt(n()))
```

@@

fig2<-ggplot(eyetracking.figure.2.data %>% 
  mutate(compatibility = factor(compatibility, 
                                levels = c("modifier", "equibiased", "instrument" ), 
                                labels = c("Modifier", "Equi-biased", "Instrument" ))), 
  aes(x=t.window, y=M, ymin=M-SE, ymax=M+SE, color=compatibility, fill=compatibility, linetype=object_type))+
  geom_ribbon(color=NA, alpha=0.3)+
  geom_line(size=1)+
  scale_color_manual(values = c( "#377eb8", "#e41a1c","#4daf4a"))+
  scale_fill_manual(values = c("#377eb8", "#e41a1c", "#4daf4a"))+
  scale_linetype(labels = c("Animal", "Instrument") )+
  theme_classic() +
  geom_vline(xintercept = animal.onset + 200) + 
  geom_vline(xintercept = instrument.onset + 200)+
  labs(y = "Proportion of looks", x = "Time relative to verb onset (ms)")+
  guides(color = guide_legend("Verb bias"),fill = guide_legend("Verb bias"), linetype = guide_legend("Gaze location"))
fig2
```
```{r, eval=FALSE}
saveRDS(fig2, "output/ETfig-in-lab.rds")
ggsave("output/ETfig-in-lab.png", plot = fig2, height = 4, width = 6)
```

```{r E4-compute-proportion-gaze-data-by-trial}
#Summarize fixations on target and instrument
eyetracking.window.summary.by.trial <- eyetracking.data.with.time.windows %>%
  group_by(unique_id, trialID, sound, compatibility, time.window) %>%
  summarize(prop.fixations.animal = sum(in.roi.animal) / n(),
            prop.fixations.instrument = sum(in.roi.instrument) / n()) %>%
  mutate(compatibility = factor(compatibility))
```

@@

model.time.window.1 <- lmer(prop.fixations.animal ~ compatibility + (1  | unique_id) + (1 | trialID), data=data.time.window.1,
                            control = lmerControl(optimizer = "bobyqa",
                                                  optCtrl = list(maxfun = 2e6)))

@@

model.time.window.2 <- lmer(prop.fixations.animal ~ compatibility + (1  | unique_id) + (1 | trialID), data=data.time.window.2)
```

@@

model.time.window.3 <- lmer(prop.fixations.animal ~ compatibility + (1 | unique_id) + (1 | trialID), data=data.time.window.3)
```
```{r E4-ET-analysis-tables}
E4_ET_model1_tab = broom.mixed::tidy(model.time.window.1) 
E4_ET_model2_tab = broom.mixed::tidy(model.time.window.2) 
E4_ET_model3_tab = broom.mixed::tidy(model.time.window.3)
